---
title       : How to become financially independent
subtitle    : 
author      : Sébastien Drouyer
job         : Computer science engineer
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## What is financial independence

* Financial independence is generally used to describe the state of having sufficient personal wealth to live, without having to work actively for basic necessities. [[`Wikipedia`](https://en.wikipedia.org/wiki/Financial_independence)]


* It means you earn more from your assets (investments) than you need to spend.

* Though it helps, you don't have to get a huge salary to become financially independent.


* For instance, if Bob need to spend 1000$ each month, and receive 1100$ each month from his investments, he is financially independent. If Steve spends 200,000$ each month but receive 100,000 from his investments, he is not financially independent.

--- .class #id 

## Why achieve financial independence

There can be many reasons:

* Freedom: you don't have to work anymore in order to live. Once you achieve financial independence, you can keep working, but you can also start a business, follow your passions, or go into an indefinite vacations...

* Security: if anything should happen, you have a solid safety net that will prevent you or you family to end up in the street.

* Retirement: of course, this is an obvious objective.

* Confidence and personal satisfaction: by knowing that you have a solid safety net and achieved a difficult objective.

--- .class #id 

## How to achieve it?

You can change many parameters:

* Spend less and save: saving is a crucial if you want to achieve financial independence. It can mean living in a smaller house than you can afford to for instance. For some people it can be difficult, but you generally don't notice it after a while.

* Earn more and save: if you can't spend less, you need to get an alternate source of income. You can start a small business or work additional hours if necessary.

* Invest: Investing your saved money is important if you want to receive earnings from it. Many different solutions (such as bonds or stocks) exist.

--- .class #id 
## An example

```{r, echo=FALSE}
library(ggplot2)
simulation_plot <- function(input) {
        salaries <- matrix(input$initial_salary, input$simulation_length, 1)
        revenue_from_savings <- matrix(0, input$simulation_length, 1)
        savings  <- matrix(input$initial_savings, input$simulation_length, 1)
        financially_independent = -1 # expense rate
        savings_revenue_superior_to_salary = -1
        for (i in 1:input$simulation_length) {
                salaries[i,1] = input$initial_salary * (1.0 + (input$estimated_raise_per_year / 100.0)) ^ (i - 1)
                salaries[i,1] = salaries[i,1] * rnorm(1, mean=1, sd=(input$salary_variance/100.0))
                if (i > 1) {
                        this_year_return_rate <- rnorm(1, mean= (input$investment_returns / 100.0), sd=(input$investment_returns_variance/100.0))
                        revenue_from_savings[i, 1] = savings[i - 1] * this_year_return_rate
                        if (revenue_from_savings[i, 1] > salaries[i,1] && savings_revenue_superior_to_salary == -1) {
                                savings_revenue_superior_to_salary = i
                        }
                        if ((revenue_from_savings[i, 1] > salaries[i - 1, 1] * (100.0 - input$salary_saving_rate) / 100.0) && financially_independent == -1) {
                                financially_independent = i
                        }
                        
                        savings[i, 1] = savings[i - 1] + revenue_from_savings[i, 1] + salaries[i - 1, 1] * input$salary_saving_rate / 100.0
                }
        }
        
        salaries <- data.frame(year=1:input$simulation_length, value=salaries)
        savings <- data.frame(year=1:input$simulation_length, value=savings)
        revenue_from_savings <- data.frame(year=1:input$simulation_length, value=revenue_from_savings)
        p <- ggplot() + 
                geom_line(data = salaries, aes(x = year, y = value, color = "Salary")) +
                geom_line(data = savings, aes(x = year, y = value, color = "Savings"))  +
                geom_line(data = revenue_from_savings, aes(x = year, y = value, color = "Returns"))  +
                xlab('Year') +
                ylab('Money') +
                labs(color="Legend")
        if (financially_independent > -1) {
                p <- p + geom_vline(xintercept = financially_independent, linetype = "longdash")
        }
        if (savings_revenue_superior_to_salary > -1) {
                p <- p + geom_vline(xintercept = savings_revenue_superior_to_salary)
        }
        
        return (p)
}

input = c()
input$simulation_length = 30
input$initial_savings = 0
input$estimated_raise_per_year = 2
input$salary_variance = 0
input$investment_returns_variance = 0
input$initial_salary = 60000
input$salary_saving_rate = 50
input$investment_returns = 8
```
If someone with a 60000$ per year salary and no initial savings saves 50% of his/her salary and invest it into a 8% per year investment (S&P500 is around 11%), then he/she will achieve financial independence in 12 years.
Of course, such an investment would have some variance; try my application to get a more detailed simulation.
```{r, fig.height=4, fig.width=7, fig.align='center'}
print(simulation_plot(input))
```
